
%%% From Proposal
Formal verification is a key to secure and reliable software, and  Operating System (OS) kernels and hypervisors form the backbone of every safety-critical software system in the world. 
Hence it is highly desirable to formally verify the correctness of these programs.

Recent work on \selfour~\cite{klein2009sel4,klein14} has shown that it is feasible to formally prove the functional correctness property of a general-purpose microkernel, but the cost of such verification is still quite prohibitive. 
It took the \selfour\ team more than 11 person years (effort for tool development excluded) to verify 7500 lines of sequential C code, yet the resulting kernel still contains 1200 lines of additional C code and 600 lines of assembly code that are not verified. 
Worse still, even after all these efforts, the current verified \selfour\ kernel cannot be used to reason about user-level programs as it does not verify important features such as virtual-memory page faults and address translation.

There are many reasons that make hard to verify the OS kernels formally.
First, OS kernels are complex artifacts; they contain many interdependent components that are difficult to untangle.
Their invariants can involve machine level details (e.g., how the virtual memory hardware works) but can also cut across multiple abstraction boundaries (e.g., different views of an address space under kernel/user or host/guest modes).
Several researchers~\cite{baumann12,vaynberg12} observed that even writing down a good and easy-to-maintain formal specification alone is already a major roadblock for any such verification effort.

Second, OS kernels are often written in C, which only supports limited forms of abstraction.  Verification of C programs is especially hard if they manipulate low-level data structures (e.g., thread queues, allocation tables).  
The \selfour\ effort used an intermediate executable specification (derived from a Haskell prototype) to hide some messy C specifics, but this alone is not enough for enforcing abstraction among different kernel components; \selfour\ had to introduce capabilities which add significant implementation complexities to the kernel.

Third, OS kernels are developed for managing and multiplexing hardware, so it is important to have a machine model that can describe hardware details.
The C language (especially ANSI C) is too high level for this purpose. For example, while most kernel code can be written in C, many key kernel concepts (e.g., context switches, address translation, page fault handling) can only be given accurate semantics at the assembly level. Consequently, we need a formal assembly model to define many kernel behaviors, but we also want to verify most kernel code at a much higher abstraction level.

Fourth, OS kernel verification would not scale if it does not  support extensibility.
One advantage of a verified kernel is the existence of formal specifications for all of its components. 
In theory, this would allow us to add certified kernel plug-ins as long as they do not violate any existing kernel invariants.
In practice, however, if we are unable to decompose kernel invariants into small independent pieces, even modifying an existing (or adding a new) verified component may force us to rewrite the proofs for the entire kernel.

Hence, OS kernel and hypervisor verification needs a novel compositional approach that can handle all of the above challenges successfully. 
Previous work~\cite{dscal15} in our group presents a verified OS kernel and hypervisor with solving those problems. 
However, there are still limitations in two verified kernels~\cite{dscal15, klein2009sel4}.
For example, both of them only support single processor systems.



%%%% From MCSLock

\section{MCS Lock}

The MCS algorithm for scalable fair inter-CPU mutex locks makes for an interesting case study in program verification.
Although the program is short, the proof is challenging.
First, the implementation of a lock algorithm can not itself use locks, so it has to rely solely on atomic memory instructions and be robust against any possible interleavings between CPUs. This is the most challenging type of concurrency, so-called lock-free programming.
Second, unlike algorithms which only promise mutual exclusion, the MCS algorithm also aims for fairness among CPUs. To check that it got it right, our correctness theorem needs to guarantee not only mutual exclusion (a safety property) but also bounded waiting time (a liveness property).

Previous work~\cite{liang:lili,ogata:mcs-lock} has studied the
correctness of the algorithm itself, but those verification efforts
did not produce executable code, and did not explore how to integrate
the proof of the algorithm into a larger system. We have created a
fully verified implementation and added it as part of the CertiKOS
kernel~\cite{certikos16}, which consists of 6500 lines of C and
assembly implementation and 135K lines of Coq proofs.

In order to manage such a large verification effort, the CertiKOS team developed a methodology known as \emph{certified (concurrent) abstraction layers}, as well as a set of libraries and theorems to support it. Previous papers~\cite{dscal15,ccal16}
described this framework, but many readers found them  dense and hard to follow because they immediately present the formalism at its most abstract and general.
This paper aims to be a complement: by zooming in on the implementation of one small part of the kernel (the MCS Lock module), we illustrate  what it is like to \emph{use} the framework, how to write specifications in the ``layers'' style, and what the corresponding proof obligations are. We hope this paper will be an easier entry point for understanding our verification framework.

As we will see, CertiKOS-style verification has several distinctive features which stem from the requirements of a large kernel. First, it is suitable for {\bf dealing with low-level code}. To make the proofs tractable we mainly work at the C level (relying on the CompCert verified compiler~\cite{Leroy-Compcert-CACM}), but sometimes we need to go lower. For example the MCS algorithm needs to use atomic CPU instructions ({\em fetch-and-store} and {\em compare-and-swap}), so we need a way to  mix C and assembly code, while stating precisely what semantics we assume that the assembly code has. At the same time, C itself is too low-level to conveniently reason about, so we need {\bf data abstraction} to hide the details about representation in memory.

Second, in order to handle large developments we need {\bf separation of responsibilities}. In a small proof of an algorithm in isolation, you can state the specification as a single pre- and post-condition which specifies the shape and ownership of the data structure, the invariants (e.g. mutual exclusion), the liveness conditions, and even the behavior of the lock's client code (the critical section code). But such a proof is not modular and not re-usable. In our development, these are done as separate refinement steps, in separate modules with explicit interfaces, and can even be the responsibility of different software developers. 

Finally, the layers approach is {\bf general purpose}, in the sense
that the same semantic framework can be used for proving all kinds of
properties. The model of program execution exposed to the programmer
is simple, mostly the same as for sequential code and with a notion of logs of events to model concurrency.
Unlike working in a special-purpose program logic, we 
did not have to add any features to show a liveness property, because we can directly reason in Coq about \textbf{how long} an execution will take. 

In the remaining parts of the paper, we first explain the C code that we will be verifying (Sec.~\ref{sec:overview}).
Then in the bulk of the paper, we explain our proof strategy by going through each abstraction layer in turn, concluding with the safety and starvation freedom properties (Sec.~\ref{sec:verification}). Finally we explain how our proofs fit as a part of the larger CertiKOS development (Sec.~\ref{sec:evaluation}) and discuss related and future work (Sec.~\ref{sec:related}).

We revisit several points that have been mentioned in previous publications:

\begin{itemize}
\item We show how to customize the machine model by adding a trusted specification of particular instructions that we need. (Sec.~\ref{subsec:lowestmachinemodel}.)

\item We locally verify the execution of a single CPU, and treat the rest of the system as an opaque \emph{concurrent context}. (Sec.~\ref{subsec:abstractoperationlayer}.)

\item We illustrate how to abstract away from a C implementation, by refining it into a functional specification which can be conveniently reasoned about. (Sec.~\ref{subsec:atomicoperation}.)

\item Similarly, we show how the same type of refinement can be used to gradually add ghost state to a specification while hiding un-needed details. (Sec.~\ref{sec:representation-ghost}.)

\end{itemize}

We also make novel contributions:


\begin{itemize}
\item It provides a concrete example of CertiKOS-style verification; in particular we can see how to customize the machine model (Sec.~\ref{subsec:lowestmachinemodel}) and how to split the verification effort into CPU-local reasoning (Sec.~\ref{subsec:eventlogandoracle} and \ref{subsec:abstractoperationlayer}).
  
\item We show a way to prove that an atomic specification refines a concurrent implementation, while still using downward rather than upward simulations. The trick is to provide a \emph{function} from low-level to high-level logs of events. (Sec.~\ref{sec:liveness-atomicity}--\ref{sec:downwards-to-upwards}.)

\item We propose a new way to specify the desired---atomic---behavior of the lock/unlock methods. To ensure liveness, the specification of the lock method itself includes a promise to later call unlock; we do this using a bounding counter. (Sec.~\ref{sec:liveness-atomicity}.)

\item And of course, we provide the first implementation of the MCS algorithm that has been both rigorously verified (with a mechanized proof) and at the same time realized (as part of a running kernel).
\end{itemize}



%
%
%\section{Key Contributions}|
%\label{chapter:introduction:sec:key-contributions}