
\section{Multi-Paxos with Witness}
\label{chapter:witnesspassing:sec:multipaxos-with-witness}


This section describes how we modeled and verified multi-Paxos using 
our approach. We present the systems specific specifications and how the
specification connects to the template to verify linearizability and soundness
of the leader. 

\subsubsection{Data type}

Concrete data types in Figure~\ref{fig:chapter:witnesspassing:basic-state} are instantiated:
\begin{itemize}
\item $\nodeid$ and $\termnum$ are $\mathbb{Z}$ type values.
\item $\seqnum$ is a pair of two values in $\mathbb{Z}$ type. The first value represents 
the \textit{index} in the multi-Paxos log that is being written, and the 
second represents the last \textit{committed index} where all log entires upto this
index contain immuatable states chosen by the acceptors.
\item $\dsvalue$ is an array of $\mathbb{Z}$ type values.
\end{itemize}

\subsubsection{Leader election}
To model the multi-Paxos leader election function based on $\ldrfunction$,
two protocol specific functions in the requester side are instantiated.
\begin{itemize}
\item $g_{pre} (l)$ adds a ghost packet that adds 1 to the term number.
\begin{center}
$g_{pre} (l) := nid.[\mbox{\textbf{inc}} \ \replay(l)[nid].rn]::l$
\end{center}
\item $g_{post} (l)$ collects acknowledgement messages and checks for the
	quorum.
	\begin{center}
$g_{post} (l) := \mbox{\textbf{let}}\ pkts := filter_{pkt}(\replay(l)[nid].rn, nid)\ \mbox{\textbf{in}} \ cal\_qrm(pkts)$
\end{center}
\end{itemize}

\subsubsection{Commit}
Similar to the leader election function, the protocol specific functions in the requester side are instantiated. 
\begin{itemize}[leftmargin=*]
\item $g_{pre} (l, u)$ is an empty function, which does nothing.
\item $g_{post} (l, u)$ collects acknowledgement messages, checks for the
	quorum, and updates the local status by applying the update function
		$u$.
\begin{center}
$
\begin{array}{lll}
 g_{post} (l, u) &:=&  \mbox{\textbf{let}}\ pkts := filter_{pkt}(\replay(l)[nid].rn, nid) \ \mbox{\textbf{in}}\\
& & \mbox{\textbf{if}} \ OK(cal\_qrm(pkts)) \ \mbox{\textbf{then}} \ nid.[u]::l\ \mbox{\textbf{else}}\ l \\
\end{array}
$
\end{center}
\end{itemize}

\subsubsection{Update function}
The object update function is relatively simple, 
which adds index and committed index numbers by 1 and applying changes to the
state:
\begin{center}
$
\begin{array}{lll}
u (st : \dsvalue)&:=&\mbox{\textbf{match}} \ st \ \mbox{\textbf{with}}\\
& & \vert~(\_, (idx, cidx), dsval)  \Rightarrow ((idx + 1, cidx + 1), dsval[idx + 1] := uf_{\dsvalue}\ dsval[idx])\\
& &  \mbox{\textbf{end}}.\\
\end{array}
$
\end{center}
Here, $u_{\dsvalue}$ is a function that is specific to the system hosted 
by multi-Paxos.

\begin{theorem}[Linearizabilty of MultiPaxos]
Multipaxos is  linearizable.
\end{theorem}

\begin{proof}
Our implementation satisfies all the requirements in Section~\ref{chapter:witnesspassing:sec:prove-safety-with-witness}. 
Therefore, proving the linearizability of MultiPaxos is straightforward with Theorem~\ref{theorem:chapter:witnesspassing:linear-spec}.
\end{proof}
